---
title: "Activity Monitoring"
output: html_document
---
## Analysis of Steps per Day

### Libraries included
*readr* for importing data set
*dplyr* for pipe and filter
*ggplot2* for time series graph

```{r}
library(readr)
library(dplyr)
library(ggplot2)
```
### The data set
This data set has data for the steps taken over the dates from October 1 
2012 to November 11 2012. The data is first imported, then filtered to remove
any entries that have NA as the number of steps. There seems to be no reason to 
believe we can impute likely values for NA, so it's best just to exclude them.

### Importing the data set
```{r}
rawdata <- read.csv("activity.csv")
```

### Cleaning out the NA's
```{r}
cleandata <- rawdata %>% 
  filter(!is.na(steps))
```

### Addtional alteration
The original dataset contains thousands of records that represent the number of
steps in five minute intervals. The assignment calls for the average number of 
steps per DAY. Therefore the data will need to be grouped by date. Since later
parts of the assignment also call for the day of the week, we will also add
a column for weekday in this step.

```{r}
cleandata <- cleandata %>% 
  group_by(date) %>% 
  summarize(steps=sum(steps)) %>% 
  mutate(weekday = weekdays(as.Date(date)))
```

### Histogram of the steps.
A histogram reveals the spread of the number of steps by day.
```{r}
hist(cleandata$steps)
```

### Mean steps per day
```{r}
mean(cleandata$steps)
```

### Median steps per day
```{r}
median(cleandata$steps)
```

### Time series plot of average steps taken.
```{r}
timeplot <- ggplot (data=cleandata) +
  aes (x=date, y=steps) +
  geom_point()

timeplot
```

